\LARGE \bf ``Those who can't do, teach'': a Teachable Robotic Agent for Children Learning Handwriting

### 2014-09-08 main.tex
Tuning, refining, improving

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

@@

Presented is a teachable humanoid agent in the context of handwriting skills
acquisition. It is hypothesised that such a system could not just engage an
unmotivated student, but could also present the opportunity for children to
experience the typical benefits experienced during human-led handwriting
interventions, such as motor mimicry. This system allows for exploring the
potential for the \emph{learning by teaching} paradigm to be employed in the
interaction, so as to stimulate meta-cognition, empathy and increased
self-esteem in the child user.

By leveraging simulated handwriting on a synchronised tablet display, a Nao
humanoid robot with limited fine motor capabilities has been configured as a
suitably embodied handwriting partner. Shape models derived from principal
component analysis of a dataset of letter trajectories are generated and allow
the robot to draw purposefully deformed letters. Learning how to write
characters well is then achieved by incorporating feedback from user
demonstrations, which allows the system to learn the optimal parameters for the
appropriate shape models.

An in situ study has been conducted with primary school classes to obtain
insight into children's use of the novel system. 
%The technical components of the system have been validated, in that children 
Children aged 6-8 were successfully 1) led to believe that the robot was
actually writing on the tablet, and 2) able to improve the robot's writing to a
level which they were satisfied with. Furthermore, the system represents a
significant step towards an innovative use for robotics which addresses a
widespread and socially meaningful challenge in education, and provides context
for robotics in education beyond the well-explored use of teaching coding- and
robotics-related subjects.

@@

Handwriting difficulties in children at an early age often negatively affect
the academic performance of the students \cite{Christensen2005}, in addition to
their self-esteem being adversely affected \cite{Malloy1995}, causing them to
shy away from expressing what they know \cite{Medwell2008}.
%A conclusion drawn
%in \cite{Hoy2011} is that any handwriting intervention studies considered in the
%systematic review which involved fewer than two practice sessions per week and
%fewer than a total of 20 practice sessions, including homework, were found to
%demonstrate ineffective results. This highlights the necessity to engage
%students in an interaction that will be sustainable over the long-term.

Successful interventions for children with handwriting difficulties involve the
student in many sessions where they are engaged in physically practising the
skill \cite{Hoy2011}. However, the link between handwriting difficulties and low
self-efficacy \cite{Engel-Yeger2009} results in children who are unmotivated to
participate in such sessions, potentially leading to a developmental arrest in
the acquisition of the skill.

@@

The \emph{learning by teaching} paradigm, which engages the target student in
the act of teaching another, has been shown to produce motivational,
meta-cognitive, and educational benefits in a range of disciplines
\cite{Rohrbeck2003}. An application which is yet to be explored, however, is the
application of learning by teaching to handwriting intervention. One reason why
this has not been investigated to date is that in order to engage a child in the
learning by teaching paradigm, there must be an appropriately unskilled peer for
them to tutor. In the context where the target child is the lowest performer in
their class, this may be a constraint. In some cases it may be appropriate for a
peer or teacher to simulate a na\"ive learner for the target child to teach,
however for an activity where one's skill level is visually evident - as is the
case in handwriting - this acting is likely to be eventually detected. As such,
there is motivation for the use of a teachable agent which can be configured for
a variety of skill levels, and for which children do not have preconceptions
about its handwriting ability.

Presented is the development of the first known learning agent which is capable
of artificially making mistakes typical of children learning handwriting, and
correcting them with external feedback. Such a system therefore exhibits the potential
to engage children in the learning by teaching paradigm in the context of
handwriting.

There are three main points which have motivated the use of a robot as the
teachable agent in the presented interaction:

\begin{enumerate}
    \item the engagement and ``protégé effect'' elicited by teaching an embodied
        agent is anticipated to be stronger than that associated with a
        tablet-based application, which children are increasingly recognising as
        disposable technology,

    \item observing a teachable agent respond to what it has been taught
        \cite{Okita2006} is a significant step in the learning by teaching
        paradigm, and therefore it may be significant for the agent to respond
        with a writing \emph{process}, rather than simply a product, and

@@

Consequently, the learning agent developed has been embodied in a commercially
affordable Nao humanoid robot. Its writing capabilities have been implemented by
leveraging synchronisation with a tablet display to simulate fine motor skills
otherwise infeasible for the robot.

Section \ref{sec:learningAlgorithm} presents the novel work in the area of
artificial intelligence to develop a learning algorithm suitable for a learning
agent in the context of handwriting.  Section \ref{sec:robotWriting} details the
extension of this algorithm to an embodied robotic learning agent, including the
novel approach for achieving simulated fine motor skills on commercially
affordable humanoid robots.  Section \ref{sec:experiment} explores the novel
contributions made to the study of human-robot interaction, in discussing the
use of the system with primary school children and its potential as a tool for
addressing wider pedagogical research questions in education.

@@

Because of the objective to use an artificial intelligence algorithm in a
reinforcement learning framework, a parameterisation of letters and their
deformities has been sought. Depending on the parameters input to the letter
models, different quality shapes can be generated. The system can therefore be
configured to improve its writing by modifying the parameters used to generate
the letters, based on feedback from the reinforcement learning partner. To this
end, a method which can generate a model from real-world data has been employed
in order to capture realistic variations in letters.

@@

One approach for a shape model which can appropriately represent realistic
variations in shapes with changes to its parameters is statistical shape
modelling. Statistical shape modelling is an application of principle component
analysis (PCA), where a linear transform which de-correlates data vectors is
found \cite{Stegmann2002} and allows for dimensionality reduction.

@@

The projection from the original $2n$-dimensional feature space $(n=70)$ to an
$N$-dimensional space is found as in (\ref{eq:projection}), where ${\bf p}$
contains the coordinates in the $N$ dimensional space with ${\bf 0}$-origin,
calculated as in (\ref{eq:paramCalc}). The $2n\times N$ matrix ${\bf \Phi }$ is
an orthogonal matrix composed of the eigenvectors ${\bf v}_i$ corresponding to
the largest $N$ eigenvalues ($\lambda_i$) of the covariance matrix of the
observations, as shown in (\ref{eq:transformCalc}) \cite{Stegmann2002}. As such,
each shape may be approximated by the mean shape plus a sum of the top $N$
eigenvectors, weighted by the parameter vector $p$.

@@

Clusters within paths of a particular letter have been identified using K-means
clustering to group different styles of writing the letter (\emph{allographs}),
such that the parameters in the resulting model are only representative of the
specific allograph of the letter in the dataset. 
% * <deanna.m.hood@gmail.com> 2014-09-03T13:05:23.576Z:
%
% in the end this was pointless so maybe not worth mentioning
%
PCA has been performed on all of the paths of a particular allograph in the
dataset at a time, to reduce the $2n$ dimensional space to one of a desired
number of dimensions.

(\ref{eq:projection}) may be used
to generate new shapes based on the parameters ${\bf p}$ which are used: ${\bf
p}={\bf 0}$ will yield the mean shape, and variations to each of the $N$ values
in ${\bf p}$ will cause a change in the shape represented by the corresponding
eigenvector (Figure \ref{fig:deviations_sPrint}).

@@

In \cite{Chandra2013}, it was found that children aged 4-6 years participating
in a handwriting peer tutoring pilot study most often made mistakes classified
as inappropriate \emph{internal proportions} or \emph{global deformations}. It
would be reasonable to consider the shapes in Figure \ref{fig:sampleLetters} in
the same general categories. Using a database of children's letters when
available may yield potential for parameterising other mistakes made by
children, such as merging strokes or decomposing a shape into multiple subparts
inappropriately. Nevertheless, sufficient capabilities for the shape model to
generate `poor' letters which are appropriate for simulating a na\"ive learning
in the context of handwriting have been demonstrated, using the shape model
generated with PCA from a dataset of only adults' writing.

@@

\begin{equation}\label{eq:param}
{\bf p}_{demo} ={\bf \Phi}^T ({\bf{x}_{demo}-\bar{x}})
\end{equation}

@@

For the statistical shape model employed by the system, the method for
responding to user demonstrations is to move the learning algorithm's parameters
towards those of the demonstration to some extent. In the results presented in
this work, \[{\bf p}^{(k+1)} = {\bf p}^{(k)} + ({\bf p}_{demo}-{\bf
p}^{(k)})\times\alpha\] has been used, where ${\bf p}$ is the parameter vector
at time step $k$, and $\alpha$ is the learning rate, between 0 and 1.  It is
possible that parameters ${\bf p}^{(k)}$ and ${\bf p}_{demo}$, which
individually yield acceptable shapes, may produce parameters ${\bf p}^{(k+1)}$
which yield an unacceptable shape. The proposed method for adapting to user
demonstration would not get stuck at such poor shapes, however: with further
demonstrations of the same letter, the system would eventually approach it. As a
result, this event would not limit the interaction proposed in Section
\ref{sec:experiment}, but it may influence the user's perception of the learning
agent. It remains to be seen if it is necessary to avoid such an event
mathematically.

@@

Using simulated handwriting provides an opportunity for the robot's writing to
appear smoother than would be achievable with a writing instrument. However, the
robot's motions must still sufficiently match the displayed trajectory in order
capture the engagement of the child user in the action.

Aldebaran's NaoQi API has been used for the inverse kinematics of the trajectory
following. The Robot Operating System (ROS)\footnote{The ROS stack for Nao is
available at: \url{http://wiki.ros.org/nao_robot}} has been used for integration
of the Nao with external reference frames using the $tf$ transformation library
\cite{Foote2013}, such as the tablet's location.

When using simulated handwriting, it is no longer necessary that the robot
engages in the typical style of handwriting with a writing instrument at a desk.
By instead having the robot point at a vertical writing surface to cause the
trajectory to appear (as in Figure \ref{fig:naoWriting}), several advantages are
presented:

@@

    \item Perhaps most significantly, the accuracy of the matching of the
        robot's motion with the trajectory displayed on the tablet is not as
        critical, because the finger is not expected to touch the tablet exactly
        at the trajectory point, in contrast to the pen tip.
        % This means that orientation degrees of freedom can be freed without significant adverse effects.

@@

\begin{figure}[thpb]
     \begin{center}
        %\subfigure{
        %    \includegraphics[width=0.4\textwidth]{figures/naoWriting1.png}
        %}%
        \subfigure{
            \includegraphics[width=0.3\textwidth]{figures/naoWriting2.png}
        }%
%
    \end{center}
    \caption{A demonstration of the robot simulating the writing of a word
        with its finger. The motion of the robot is synchronised with the display of the tablet, communicating over ROS.
     }%

@@

Therefore, the system presented utilises the context where the robot is
simulating handwriting by pointing at the tablet\footnote{Teachers interviewed
for their feedback on the system advised that children are asked to draw letters
in the air in a similar manner as part of their handwriting education.}. Because
motion planning is performed with respect to the hand of the robot, rather than
its fingertip, the one or two of the orientation degrees of freedom of the hand
are fixed to keep the finger approximately perpendicular to the writing surface,
depending on the desired accuracy. The remaining free orientation(s), coupled
with the whole-body motion control available, allow for a sufficient working
space for writing on the entire tablet.

@@

To instruct the robot where to write, the robot has been configured to detect a
particular fiducial marker, a \emph{chilitag}\footnote{See
\url{https://github.com/chili-epfl/ros_markers} for more information on the
fiducial markers used.}, with the camera located in its head and use that as the
origin of the writing surface (Figure \ref{fig:tabletDetection}). When used in
an interaction with a child, this allows the user to move the tablet as required
for the interaction. As publishing the robot's camera has been identified as a
computationally expensive task which can disrupt the robot's timing of
synchronised writing, it is only done when the robot is not writing, and the
tablet is assumed to be stationary during the writing process.

@@

The fusion of the embodied handwriting agent developed with the handwriting
learning algorithm presented in Section \ref{sec:learningAlgorithm} involves the
integration of three components. The robot and Android tablet application
present the writing process/result to the user, as in Section
\ref{sec:tabletSynch}. Furthermore, the tablet application is extended to act as
the primary medium for capturing user interaction, and publishes the user's
demonstrations when user is satisfied with their writing. The user
demonstrations are received by an interaction controller, running on a desktop
computer. It is responsible for getting the Nao to prompt and respond
appropriately to feedback received using a finite state machine to manage the
interaction stage. It also manages the learning algorithm, by inferring the
letter which user demonstrations are for, based on their position on the tablet,
and publish the shapes to be written which are generated by the learning
algorithm in response to the feedback.

@@

In addition to constituting a technically novel system, the teachable robotic
agent presented represents a tool which may be used for investigating social and
pedagogical research questions. One such question that is being worked towards
is what impact to the outcomes of a handwriting intervention the addition of
such a teachable robotic agent would have.

%\subsection{Interaction Context}

The interaction context which has been developed as a framework for addressing
such questions is one where participants are asked to improve the robot's
handwriting by correcting its letters. Figure~\ref{fig:pilotInteraction}
illustrates an example interaction sequence between the participant and the
robot, which consists of the following stages.

@@

    \item The participant shows the robot one of five different 3-letter words
        to write, which, combined, consist of 7 different letters (`c', `e',
        `n', `o', `s', `u', `w'). Fiducial markers placed on the word cards to
        allow for detection with a camera.

    \item The robot responds to the word request verbally and writes the letters
        by pointing at the tablet - placed in front of it with a vertical
        orientation - and imitating tracing the letters as they appear
        simultaneously by being displayed by the tablet.

    \item The robot asks for feedback from the participant and they demonstrate
        how to write the letter which they feel needs to be corrected. The
        demonstration is performed by drawing on the tablet with a stylus,
        either with the tablet still in a vertical orientation or by unfolding
        the stand and laying the tablet flat on the desk. The position of the
        participant's demonstration on the tablet encodes the shape letter that
        it is a demonstration for. The participant can clear their shapes if
        they are unhappy with them or they are in the wrong position. When the
        button to signal that it is the robot's turn is pressed, the
        demonstration is sent to the controller.

    \item The interaction iterates, with participants taking turns to interact
        with the robot.

@@

\begin{figure}[thpb!]
     \begin{center}
%
        \subfigure[The user shows a card to the robot with a word to write.]{%
            \label{fig:first}
            \includegraphics[width=0.2\textwidth]{figures/1card.png}
        }~
        \subfigure[The robot writes the word seen on the card and asks for feedback.]{%
           \label{fig:second}
           \includegraphics[width=0.2\textwidth]{figures/2word.png}
        }\\ %  ------- End of the first row ----------------------%
        \subfigure[The user provides feedback on the letters written via demonstration.]{%
            \label{fig:third}
            \includegraphics[width=0.2\textwidth]{figures/4correct.png}
        }~
        \subfigure[The robot responds to the feedback and asks for more.]{%
            \label{fig:fourth}
            \includegraphics[width=0.2\textwidth]{figures/5iterateAsk.png}
        }%
%
    \end{center}
    \caption{%
        A user interacting with the robot in the learning by teaching interaction using demonstrations.
     }%
   \label{fig:pilotInteraction}
\end{figure}

%\subsection{Outcomes}

A pilot study has been conducted to evaluate the interaction in preparation for
an experiment to be conducted on the 10th-11th of June, 2014. The pilot study
consisted of four groups of approximately 8 children, aged 6-7 years. Outcomes
from the pilot study include information on how children interact with the
system, sample letters written by children, opinions from the teachers involved,
and insight into what experimental conditions might be appropriate for the
experiment.

The system was validated, in that the learning algorithm and robotic writing
were believed by the children, and that the system withstood the interaction for
a total of 65 minutes, with the robot writing 96 letters, 49 of which in
response to demonstrations from the children, during that time.

Another key point which came from the pilot study is that children were observed
giving advice to the child designated as the letter demonstrator, potentially
giving rise to the higher level paradigm of learning by \emph{teaching to
teach}.

%\begin{itemize}
%
%\item Children would not intuitively write over the robot's letters, but rather
%above them. When instructed to write over the robot's letter, some participants
%would trace the letter as it had previously appeared as their demonstration. As
%such, the requested demonstration positioning will be modified for the
%experiment.
%
%\item While children spoke to the robot each time it requested feedback -
%suggesting that the anthropomorphised the robot - they also laughed when it
%wrote letters which were incorrect, which may suggest that they had not
%developed an empathetic connection with the robot by that point. 
%
%\item Children were observed giving advice to the child designated as the
%letter demonstrator, potentially giving rise to the higher level paradigm of
%learning by teaching to teach. 
%
%\end{itemize}

As a consequence of this observation, the upcoming experiment will address the
effect of the number of children interacting with the robot in order to
investigate the potential of the system to not only be used to set the context
of children tutoring the robot, but also children interacting as peer tutors
with each other.

%\section{ADDITIONAL REQUIREMENTS}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% * <deanna.m.hood@gmail.com> 2014-09-05T11:44:33.364Z:
%
% capitalise inside the phrase?
%

% * <deanna.m.hood@gmail.com> 2014-09-05T11:44:54.921Z:
%
% Ana Paiva is a reviewer....
%

@@

Presented is a robotic partner which children can teach handwriting. It is
hypothesised that such a system could not just engage an unmotivated student,
but could also present the opportunity for children to experience the typical
benefits encountered during human-led handwriting interventions, such as motor
mimicry. Moreover, this system allows for exploring the potential for the
\emph{learning by teaching} paradigm to be employed in the interaction, so as to
stimulate meta-cognition, empathy and increased self-esteem in the child user.

By leveraging simulated handwriting on a synchronised tablet display, a Nao
humanoid robot with limited fine motor capabilities has been configured as a
suitably embodied handwriting partner. Shape models derived from principal
component analysis of a dataset of letter trajectories are generated, and allow
the robot to draw purposefully deformed letters. Learning how to write
characters well is then achieved by incorporating feedback from user
demonstrations, which allows the system to learn the optimal parameters for the
appropriate shape models.

Preliminary in situ studies have been conducted with primary school classes to obtain
insight into children's use of the novel system. 
%The technical components of the system have been validated, in that children 
Children aged 6-8 were successfully 1) led to believe that the robot was
actually writing on the tablet, and 2) able to improve the robot's writing to a
level which they were satisfied with. Furthermore, the system represents a
significant step towards an innovative use for robotics which addresses a
widespread and socially meaningful challenge in education, and provides context
for robotics in education beyond the well-explored use of teaching coding- and
robotics-related subjects.

@@

Handwriting difficulties in children at an early age often negatively affect
the academic performance of the students \cite{Christensen2005}, in addition to
their self-esteem being adversely affected \cite{Malloy1995}, causing them to
shy away from expressing what they know \cite{Medwell2008}.
%A conclusion drawn
%in \cite{Hoy2011} is that any handwriting intervention studies considered in the
%systematic review which involved fewer than two practice sessions per week and
%fewer than a total of 20 practice sessions, including homework, were found to
%demonstrate ineffective results. This highlights the necessity to engage
%students in an interaction that will be sustainable over the long-term. 
Successful interventions for children with handwriting difficulties involve the
student in many sessions where they are engaged in physically practising the
skill \cite{Hoy2011}. However, the link between handwriting difficulties and low
self-efficacy \cite{Engel-Yeger2009} results in children who are unmotivated to
participate in such sessions, potentially leading to a developmental arrest in
the acquisition of the skill.

@@

The \emph{learning by teaching} paradigm, which engages the target student in
the act of teaching another, has been shown to produce motivational,
meta-cognitive, and educational benefits in a range of disciplines
\cite{Rohrbeck2003}. An application which is yet to be explored, however, is
that of handwriting intervention. One reason why this has not been investigated
to date may be that in order to engage a child in the learning by teaching
paradigm, there must be an appropriately unskilled peer for them to tutor. In
the context where the target child is the lowest performer in their class, this
may present a logistical constraint. In some cases, it may be appropriate for a
peer or teacher to simulate a na\"ive learner for the target child to teach,
however for an activity where one's skill level is visually evident - as is the
case in handwriting - this acting is likely to be eventually detected. As such,
there is motivation for the use of a teachable agent which can be configured for
a variety of skill levels, and for which children do not have preconceptions
about handwriting ability.

Presented is the development of the first known teachable agent which is capable
of artificially making mistakes typical of children learning handwriting, and
correcting them with external feedback. Such a system therefore exhibits the potential
to engage children in the learning by teaching paradigm in the context of
handwriting.

There are three main points which have motivated the use of a robot as the
teachable agent in the presented work:

\begin{enumerate}
    \item the engagement and ``protégé effect'' elicited by teaching an embodied
        agent is anticipated to be stronger than that associated with a
        tablet-based application, which children are increasingly recognising as
        disposable technology,% * <deanna.m.hood@gmail.com> 2014-09-08T11:07:10.004Z:
%
% ... citation needed?
%

    \item observing a teachable agent respond to what it has been taught
        is a significant step in the learning by teaching
        paradigm \cite{Okita2006}, and therefore it may be significant for the agent to respond
        with a writing \emph{process}, rather than simply a product, and

@@

Consequently, the learning agent developed has been embodied in a commercially
affordable Nao humanoid robot. Its writing capabilities have been implemented by
leveraging synchronisation with a tablet display to simulate fine motor skills
otherwise infeasible for the robot. 
% * <deanna.m.hood@gmail.com> 2014-09-08T11:08:36.426Z:
%
% linking sentence...
%

Section \ref{sec:learningAlgorithm} presents the novel work in the area of
artificial intelligence to develop a learning algorithm suitable for a teachable
agent in the context of handwriting.  Section \ref{sec:robotWriting} details the
extension of this algorithm to an embodied robotic learning agent, including the
novel approach for achieving simulated fine motor skills on commercially
affordable humanoid robots.  Section \ref{sec:experiment} explores the novel% * <deanna.m.hood@gmail.com> 2014-09-08T11:09:28.480Z:
%
% use a different word
%
contributions made to the study of human-robot interaction, in discussing the
use of the system with primary school children and its potential as a tool for
addressing wider pedagogical research questions in education.

@@

For the development of a reinforcement learning algorithm in the context of
handwriting, a parameterisation of letters and their deformities has been sought
to generate the necessary letter models. Depending on the parameters input to
the letter models, different quality shapes can be generated. The system can
therefore be configured to improve its writing by modifying the parameters used
to generate the letters, based on feedback from the reinforcement learning
partner. To this end, a method which can generate a model from real-world data
has been employed in order to capture realistic variations in letters.

@@

One approach for a shape model which can appropriately represent realistic
variations in shapes is statistical shape
modelling. Statistical shape modelling is an application of principle component
analysis (PCA), where a linear transform which de-correlates data vectors is
found \cite{Stegmann2002} and allows for dimensionality reduction.

@@

The projection from the original $2n$-dimensional feature space $(n=70)$ to an
$N$-dimensional space is found as in (\ref{eq:projection}), where ${\bf p}$
contains the coordinates in the $N$-dimensional space with ${\bf 0}$-origin,
calculated as in (\ref{eq:paramCalc}). The $2n\times N$ matrix ${\bf \Phi }$ is
an orthogonal matrix composed of the eigenvectors ${\bf v}_i$ corresponding to
the largest $N$ eigenvalues ($\lambda_i$) of the covariance matrix of the
observations, as shown in (\ref{eq:transformCalc}) \cite{Stegmann2002}. As such,
each shape may be approximated by the mean shape plus a sum of the top $N$
eigenvectors, weighted by the parameter vector $p$.

@@

Clusters within paths of a particular letter have been identified using K-means
clustering to group different styles of writing the letter (\emph{allographs}),
such that the parameters in the resulting model are only representative of the
specific allograph of the letter in the dataset. 
% * <deanna.m.hood@gmail.com> 2014-09-03T13:05:23.576Z:
%
% in the end this was pointless so maybe not worth mentioning
%
PCA has been performed on all of the paths of a particular allograph in the
dataset at a time, to reduce the $2n$-dimensional space to one of a desired
number of dimensions.

(\ref{eq:projection}) may also be used
to generate new shapes based on the parameters ${\bf p}$ which are used: ${\bf
p}={\bf 0}$ will yield the mean shape, and variations to each of the $N$ values
in ${\bf p}$ will cause a change in the shape represented by the corresponding
eigenvector (Figure \ref{fig:deviations_sPrint}).

@@

In \cite{Chandra2013}, it was found that children aged 4-6 years participating% * <deanna.m.hood@gmail.com> 2014-09-05T11:43:42.851Z:
%
% this is not peer-reviewed!
%
in a handwriting peer tutoring pilot study most often made mistakes classified
as inappropriate \emph{internal proportions} or \emph{global deformations}. It
would be reasonable to consider the shapes in Figure \ref{fig:sampleLetters} in
the same general categories. Using a database of children's letters when
available may yield potential for parameterising other mistakes made by
children, such as merging strokes or decomposing a shape into multiple subparts
inappropriately. Nevertheless, sufficient capabilities for the shape model to
generate `poor' letters which are appropriate for simulating a na\"ive learning
in the context of handwriting have been demonstrated, using the shape model
generated with PCA from a dataset of only adults' writing.

@@

\begin{equation}\label{eq:param}
{\bf p}_{demo} ={\bf \Phi}^T ({\bf{x}_{demo}}-\bf{\bar{x}})
\end{equation}

@@

For the statistical shape model employed by the system, the method for
responding to user demonstrations is to move the learning algorithm's parameters
towards those of the demonstration. In the results presented in this work, the
update equation shown in \ref{eq:update} has been used, where ${\bf p}$ is the
learned parameter vector at time step $k$, and $\alpha$ is the learning rate,
between 0 and 1.

\begin{equation}\label{eq:update}
{\bf p}^{(k+1)} = {\bf p}^{(k)} + ({\bf p}_{demo}-{\bf
p}^{(k)})\times\alpha
\end{equation}

It is possible that parameters ${\bf p}^{(k)}$ and ${\bf p}_{demo}$, which
individually yield acceptable shapes, may produce parameters ${\bf p}^{(k+1)}$
by which yield an unacceptable shape. The proposed method for adapting to user
demonstration would not get stuck at such poor shapes, however: with further
demonstrations of the same letter, the system would eventually approach the
demonstration shape. As a result, the event of poor-looking intermediate letters
would not limit the interaction later proposed in Section \ref{sec:experiment}
in a technical sense, but it may influence the user's perception of the learning
agent. It remains to be seen if it is necessary to avoid such an event
mathematically.

@@

Using simulated handwriting provides an opportunity for the robot's writing to
appear smoother than would be achievable with a writing instrument. However, the
robot's motions must still sufficiently match the displayed trajectory in order
capture the engagement of the child user in the action. Aldebaran's NaoQi API
has been used for the inverse kinematics of the trajectory following. The Robot
Operating System (ROS)\footnote{The ROS stack for Nao is available at:
\url{http://wiki.ros.org/nao_robot}} has been used for integration of the Nao
with external reference frames using the $tf$ transformation library
\cite{Foote2013}, such as the tablet's location.

When using simulated handwriting, it is no longer necessary that the robot
engages in the typical style of handwriting, with a writing instrument at a desk.
By instead having the robot point at a vertical writing surface to cause the
trajectory to appear (as in Figure \ref{fig:naoWriting}), several advantages are
presented:

@@

    \item Perhaps most significantly, the accuracy of the matching of the
        robot's motion with the trajectory displayed on the tablet is not as
        critical, because the finger is not expected to touch the tablet exactly
        at the trajectory point, in contrast to a pen tip.
        % This means that orientation degrees of freedom can be freed without significant adverse effects.

@@

\begin{figure}[thpb]
     \begin{center}
        %\subfigure{
        %    \includegraphics[width=0.4\textwidth]{figures/naoWriting1.png}
        %}%
        \subfigure{
            \includegraphics[width=0.3\textwidth]{figures/naoWriting2.png}
        }%
%
    \end{center}
    \caption{A demonstration of the robot simulating the writing of a word with
    its finger. The motion of the robot is synchronised with the display of the
    tablet, communicating over ROS.}%

@@

Therefore, the system presented utilises the context where the robot is
simulating handwriting by pointing at the tablet\footnote{Teachers interviewed
for their feedback on the system advised that children are asked to draw letters
in the air in a similar manner as part of their handwriting education.}. As
interacting with a tablet with one's finger is not uncommon, this may aid the
acceptance of the writing style by users.

Because
motion planning is performed with respect to the hand of the robot, rather than
its fingertip, one or two of the orientation degrees of freedom of the hand
are fixed to keep the finger approximately perpendicular to the writing surface,
depending on the desired accuracy. The remaining free orientation(s), coupled
with the whole-body motion control available, allow for a sufficient working
space for writing on the entire tablet.

@@

To instruct the robot where to write, the robot has been configured to detect a
particular fiducial marker, a \emph{chilitag}\footnote{See
\url{https://github.com/chili-epfl/ros_markers} for more information on the
fiducial markers used.}, with the camera located in its head, and use that to
determine the relative position of the writing surface (Figure
\ref{fig:tabletDetection}). When used in an interaction involving a user, this
allows a user to move the tablet as required for the interaction. As publishing
the robot's camera has been identified as a computationally expensive task which
can disrupt the robot's timing of synchronised writing, it is only done when the
robot is not writing, and the tablet is assumed to be stationary during the
writing process.

@@

The fusion of the embodied handwriting agent developed with the handwriting
learning algorithm presented in Section \ref{sec:learningAlgorithm} involves the
integration of three components: the robot, the tablet, and a centralised controller. % * <deanna.m.hood@gmail.com> 2014-09-08T12:03:16.473Z:
%
% centralised is the right word yeah? all messages pass through the controller
%
The robot and Android tablet application present the writing process/result to
the user, as in Section \ref{sec:tabletSynch}. Furthermore, the tablet
application is extended to act as the primary medium for capturing user
interaction, and publishes the user's demonstrations when user is satisfied with
their writing. The user demonstrations are received by the interaction
controller, running on a desktop computer. It is responsible for getting the Nao
to prompt and respond appropriately to feedback received, using a finite state
machine to manage the interaction stage and various system inputs. It also
manages the learning algorithm, by inferring the letter which user
demonstrations are for, based on their position on the tablet, and publishes the
shapes to be written which are generated by the learning algorithm in response
to the feedback.

@@

In addition to constituting a technically novel system, the presented teachable
robotic agent represents a tool which may be used for investigating social and
pedagogical research questions. One such question is what impact the addition of
such a teachable robotic agent would have on the outcomes of a typical
handwriting intervention.

\subsection{Interaction Context}

The interaction context which has been developed as a framework for addressing
relevant research questions is one where participants are asked to improve the robot's
handwriting by correcting its letters. Figure~\ref{fig:pilotInteraction}
illustrates an example interaction sequence between the participant and the
robot, which consists of the following stages.

@@

    \item The participant shows the robot one of seven different 3-letter words
        to write, which, combined, consist of 7 different letters (`c', `e',
        `n', `o', `s', `u', `w'). Fiducial markers which are printed on the word
        cards allow them to be detected with the robot's camera.

    \item The robot responds to the word request verbally, and writes the
        letters by pointing at the tablet and imitating tracing the letters. The
        tablet, placed in front of the robot with a vertical orientation,
        displays the letters in synchronisation with the robot's movements.

    \item The robot asks for feedback from the participant and they demonstrate
        how to write the letter which they feel needs to be corrected. The
        demonstration is performed by drawing on the tablet with a stylus. The
        tablet may be moved into the most appropriate position for the child to
        write on it. Only one letter may be demonstrated at a time, and the
        position of the participant's demonstration on the tablet encodes the
        letter that it is a demonstration for. The participant can remove and
        repeat their letter if they are unhappy with it.
        %, or if it is in the wrong position as determined by the researcher
        %conducting the interaction
        When the participant is satisfied with the demonstration, a button on
        the tablet which signals that it is the robot's turn is pressed, and the
        demonstration is sent to the controller.

    \item The controller produces an adapted letter for the robot to write in
        response to the participant's feedback, and the interaction iterates,
        with participants taking turns to interact with the robot if necessary.
        When the child is satisfied with the robot's performance on all letters,
        they may use the ``test'' card and an additional word on which they will
        verbally evaluate the robot's performance.

@@

\begin{figure}[thpb!]
     \begin{center}
%
        \subfigure[The user shows a card to the robot with a word to write.]{%
            \label{fig:first}
            \includegraphics[width=0.2\textwidth]{figures/1card.png}
        }~
        \subfigure[The robot writes the word seen on the card and asks for feedback.]{%
           \label{fig:second}
           \includegraphics[width=0.2\textwidth]{figures/2word.png}
        }\\ %  ------- End of the first row ----------------------%
        \subfigure[The user provides feedback on the letters written via demonstration.]{%
            \label{fig:third}
            \includegraphics[width=0.2\textwidth]{figures/4correct.png}
        }~
        \subfigure[The robot responds to the feedback, until the user is satisfied.]{%
            \label{fig:fourth}
            \includegraphics[width=0.2\textwidth]{figures/5iterateAsk.png}
        }%
%
    \end{center}
    \caption{%
        A user engaging with the robot in the \emph{learning by teaching}
    interaction, using demonstrations as feedback.}%

   \label{fig:pilotInteraction}
\end{figure}

\subsection{Outcomes of Preliminary Studies}% * <deanna.m.hood@gmail.com> 2014-09-05T08:43:38.105Z:
%
% I guess this is what we call the non-experiments.. Are the dates important?
%

Preliminary studies at two schools in the Geneva area have been conducted to
evaluate the feasibility and technical soundness of the interaction system
proposed. A pilot study at the first school consisted of four groups of
approximately 8 english-speaking children, aged 6-7 years. The children
interacted with the system for a total of 65 minutes, with the robot writing 96
letters, 49 of which in response to demonstrations from the children.

As a result of the pilot study, two key observations were acted on. The first
was that children appeared to have a difficult time providing demonstrations to
the robot in the same place as previously-written letters. At the time, the
system required the children to write ontop of a letter of the same type as the
one which they were demonstrating, and the children seemed to find this
counter-intuitive and would occasionally just trace the robot's letter as it
appeared. As such, the technical components of the system were extended to allow
the children the opportunity to write around previously-written letters instead
of just ontop.

The second key point which came from the pilot study was that children were
observed giving advice to the child designated as the letter demonstrator,
potentially giving rise to the higher level paradigm of learning by
\emph{teaching to teach}. As a consequence of this observation, the study which
followed at the second school was designed to further observe the effect of the
number of children interacting with the robot.
%in order to investigate the potential of the system to not only be used to set
%the context of children tutoring the robot, but also children interacting as
%peer tutors with each other.

The study at the second school involved 21 french-speaking students: 7 children
interacted with the robot by themselves, and 7 sessions included the remaining
14 students interacting with the robot in pairs. The duration of the sessions
was between 8 and 15 minutes, with an average of 11.4 minutes (standard
deviation of 2.3). % * <deanna.m.hood@gmail.com> 2014-09-05T09:25:41.865Z:
%
% We prompted them when time was short, didn't we? Or did they just say test
% whenever they wanted to...
%

The following conclusions were successfully drawn about the system as a result
of the study at the second school:

\begin{itemize}% * <deanna.m.hood@gmail.com> 2014-09-05T09:34:46.545Z:
%
% taking this out of bullet points gains extra space in paper if needed
%

    \item \textbf{The simulated robotic writing was believed by the children in all
    cases.} There were, at times, 
    % * <deanna.m.hood@gmail.com> 2014-09-05T11:59:03.640Z:
    %
    % how many
    %
    questions about the robot's writing method at the beginning of the interaction,
    but when advised that the robot ``tells the tablet what it wants to write,''
    this was accepted by the children. To allow the children an opportunity to
    express doubt in the robot's ability to write without explicitly asking them if
    the robot was pretending, the children were asking if they believed that the
    robot would be able to write its name, \emph{Nao}, which had not previously been
    demonstrated. It was reasoned that this would prompt the children to question
    the point of the robot actually \emph{writing} if they had any concerns that
    they were being fooled, but this was never the case. In the event where the
    robot's writing was not synchronised with the tablet, this did not affect the
    response of the child. If older children participate in the interaction study,
    which may be likely as children with lasting handwriting difficulties are used
    as participants, it may become more important to invest time into the
    believability of the robot's writing scheme. however for 6-8 year olds the
    proposed setup has been proven viable.
